Project-I by Group Sydney

### 2015-10-24 report/paper.tex
Finish testModels

To separate the data, we have observed that the feature $2$ and $16$ could help us. Figure \ref{fig:feature2}. We can observe $11$ misclassified data (green points), which we will remove them in order to not corrupt our model.  So, this feature can allow us to find the first cluster. For the two others clusters, we need to oberve the feature $16$. Figure \ref{fig:feature16}. We can observe $14$ misclassified data (green points and one blue). They also will be consider as outliers and remove. We set the threshold in order to minimize the number of misclassified data samples. Those thresholds are $0.42$ for the feature $2$ and $1.17$ for the feature $16$.

@@

Since our matrix is ill-conditioned, using the least squares with normal equation wasn't suitable. Moreover, the results obtained were less good (not so much) compared to the ridge regression method, and so we will not report the results obtained with this method. The method using the least squares with gradient descent was appropriated, but compared to the ridge regression, it was a little bit less efficient and was also slower to tune correctly the learning step. This is the reason why we will not report the results using this method. Therefore, we have so used the ridge regression method, which is suitable because we should lift the eigenvalues, faster and a little bit more efficient than least squares using gradient descent. Only results obtained with this latter will be reported.

@@

In order to evaluate our different models, we first have split our data ($\mathbf{X}$) in two sets of $80\%$ and $20\%$, for the training and the test sets. We learn with our models on the training set and use the test set in order to estimate the error using the \textit{Root Mean Square Error}. Moreover, we have also used k-cross validation (see next section for the values of k). We repeated the experiment $100$ times with different seed in order to split the data in a unique random way for each trial. Finally, we made varied the lambda from $10^{-5}$ to $10^{5}$ with $500$ values between.

@@

The figure XXX represents the \textit{RMSE} for each models. The first model used in the constant one, which will allow us to compare improvements with others model. It simply returns the value of the mean of the outputs and doesn't depend on the inputs. Our second model consists of using also a constant model as in first, but this time separating the data in the three clusters identified during the exploratory data analysis. We can observed a big improvement, justifying our choice to split the data in clusters.

The third model does a ridge regression on the overall data, in order to check that the separation of the data into clusters is better, even using only constant model per cluster. Hopefully, this is the case. The next model add the normalization on the input variables, another way to check if the separation of the data into cluster was a good things. Normalization is generally a good thing and often a necessity depending on the algorithms used (e.g. gradient descent ones). It improves only slightly but because normalization is a good practice, we keep it.

The fifth model consists of the separation of the data, using the normalization per cluster and finally does a ridge regression per cluster. With this manner, we can tune each cluster independently. We can see an important improvement because of the separation of the data. The next model adds the dummy encoding for the categorical variables. We can observe that it seems to be equivalent.

The seventh model is about using polynomial basis functions for each cluster. We have to be careful about the polynomial to avoid the case where $D > N$ (especially for the cluster two and three, because there are much smaller than the first one). This model doesn't use dummy encoding and so can add higher polynomial. The degrees of the polynomial for the clusters are $3,6$ and $3$. When we add higher degrees, we can directly observe overfitting because suddenly the training error decreases to a value near zero and the test errors grows exponentially, and so we know that our degrees are not to high. The eight model is similar to the seventh one, except that this time it uses dummy encoding and have lower degrees : $3,5$ and $2$, due to the number of input variables ($92$ vs $76$). We can see that these feature transformations significantly improve the prediction accuracy compared to others model.

@@

In hope to minimize the error, we tried different feature transformations. The first try was add polynomial of them self as $X = [X X^2 X^3 ...]$ for which each cluster has is own degree. We made varied the different degrees and also the lambdas in order to keep the lambda which minimize the test error.

>>>>

To separate the data, we have observed that the feature $2$ and $16$ could help us. Figure \ref{fig:feature2}. We can observe $11$ misclassified data (green points), which we will remove them in order to not corrupt our model.  So, this feature can allow us to find the first cluster. For the two others clusters, we need to oberve the feature $16$. Figure \ref{fig:feature16}. We can observe $14$ misclassified data (green points and one blue). They also will be consider as outliers and remove. We set the threshold in order to minimize the number of misclassified data samples. Those thresholds are $0.42$ for the feature $2$ and $1.17$ for the feature $16$. So the final size of our clusters are $1937, 563$ and $273$.

@@

Since our matrix is ill-conditioned, using the least squares with normal equation wasn't suitable. Moreover, the results obtained were less good (not so much) compared to the ridge regression method, and so we will not report the results obtained with this method. The method using the least squares with gradient descent was appropriated, but compared to the ridge regression, it was a little bit less efficient and was also much slower to tune correctly the learning step. This is the reason why we will not report the results using this method. Therefore, we have so used the ridge regression method, which is suitable because we should lift the eigenvalues, faster and a little bit more efficient than least squares using gradient descent. Only results obtained with this latter will be reported.

@@

In order to evaluate our different models, we first have split our data ($\mathbf{X}$) in two sets of $80\%$ and $20\%$, for the training and the test sets. We choose this percentage in order to have still enough data for the training set and also enough in order to estimate correctly our models on the testing set with a concrete size. We learn with our models only on the training set and use the test set in order to estimate the error using the \textit{Root Mean Square Error}. Moreover, we have also used k-cross validation (see next section for the values of k) to tune the parameters like lambda of the different polynomial degrees. We repeated the experiment $100$ times with different seed in order to split the data in a unique random way for each trial. Finally, we made varied the lambda from $10^{-5}$ to $10^{5}$ with $500$ values between.

@@

The figure XXX represents the \textit{RMSE} for each models. The first model used in the constant one, which will allow us to compare improvements with others model. It simply returns the mean of the outputs and doesn't depend on the inputs. Our second model does a ridge regression on the overall data using $10$ fold cross validation, because we have a lot of data, splitting them into $10$ fold seems reasonable. As expected, this is much better than previous model because we take into account the input features. Next model is similar to the second one except that we add normalization on the real inputs (it doesn't make sense to also normalize categorical variables). Normalization is generally a good thing and often a necessity depending on the algorithms used (e.g. gradient descent ones). It improves only slightly but because normalization is a good practice, we keep it.

At this stage, we want to try to separate our data into cluster. The fourth model consists of using a constant model as in first but this time separating the data into the three clusters identified during the exploratory data analysis. We can observe a big improvement, justifying our choice to split the data in clusters. The next model is about doing a ridge regression per cluster, each one having its own lambda and its own normalization. We also use k-cross validation where the first cluster has a $k_1=10$ ($1937$) because we have a lot of data, the second $k_2=6$ ($563$ samples) and the last $k_3=4$ ($273$) because it doesn't have so much data and so we have to choose a smaller k. As expected we obtained another big improvement justifying again the choice of splitting the data.

The sixth model consists of using dummy encoding for the categorical variables. As explained in the exploratory data analysis, we use $k-1$ features to represent a $k$ categorical variable. This is a common practice. Unfortunately, we didn't get major improvements.

The two next models is about using polynomial basis functions for each cluster. Categorical variables aren't concerned, they are just replicated for each degree, because it doesn't make any sense to build polynomial for them. We normalize only when we did our feature transformations, in order to avoid loosing information. We have to be careful about the polynomial to avoid the case where $D > N$ (especially for the cluster two and three, because there are much smaller than the first one). The seventh model doesn't use dummy encoding and the eight yes.

For the seventh model, the degrees of the polynomial for the clusters are $3,6$ and $3$. When we add higher degrees, we can directly observe overfitting because suddenly the training error decreases to a value near zero and the test errors grows exponentially, so with this manner we know that our degrees are not to high. The eighth model is similar to the seventh one, except that this time it uses dummy encoding and has lower degrees : $3,5$ and $2$, due to the number of input variables ($92$ vs $76$). We can see that these feature transformations significantly improve the prediction accuracy compared to others model. However, dummy encoding seems to bit a bit more efficient.

@@

In hope to minimize the error, we tried different feature transformations. The first try was add polynomial of them self as $X = [X X^2 X^3 ...]$ for which each cluster has is own degree. We made varied the different degrees and also the lambdas in order to keep the lambda which minimize the test error. We also try feature transformations where the degree is a root, but unfortunately it was worst. We also wanted to try to just pow features with different degrees (without add polynomial of themselves) but we couldn't do it due to a lack of time.
